---
title: "Analysis of the factors that contribute to the industrial conflict (A changer)"
author: "Clara le Draoulec, Maeva Kameni"
date: "March 22, 2022"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_width: 5
    fig_height: 4
    fig_caption: yes
geometry: margin=2.5cm
fontsize: 12 pt
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[R]{Clara le Draoulec, Maeva Kameni}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#install.packages('kableExtra')
library(dplyr)
library('kableExtra')
library("olsrr")
```

## Introduction:

In the United States, in order to legislate the memberships in labor union, right-to-work laws have been passed, known as the Taft–Hartley Act. However, this law is not imposed in all states. Therefore, in this analysis, we propose to address if there is a difference in terms of industrial conflict between unionized and non-unionized states and determine to what extent certain factors might contribute to this conflict.

Four factors are selected as explanatory variables. The first variable is the degree of unionization of the manufacturing labor force. The second variable is a dummy variable and describes the status of the state, where 0 represents right-to-work states and 1 is assigned to union-shop states. The third variable is a measure of the variation in the industrial structure between states. The latter aims at describing how the union strength is distributed throughout the industries of a state. Finally, the fourth variable represents the size agricultural labor force
 
```{r processing, echo = FALSE, eval=FALSE}
## Pre-processing:
#During this step, we will open the data file and make sur that every elements contained is properly #set and that the data variable is ready to be used for further analysis.
#open the data file

data <- read.table('data.rtf', header=FALSE, sep="", skip = 9)

#rename the column
names(data)[1] <- 'State_postal_abbreviation'
names(data)[2] <- 'Degree_of_conflict' #percentage of worker hours lost in stoppages
names(data)[3] <- 'Degree_of_unionization' #percent of workforce
names(data)[4] <- 'Union_shop_dummy' #1=Non-right to work 0=Right to work
names(data)[5] <- 'Sector' #percent in food, textiles and apparel
names(data)[6] <- 'Ag_force' #percent working in agriculture


#make sure the format is alright
data$Ag_force <-  gsub("\\\\", "", data$Ag_force)
data$Ag_force <- gsub("}", "", data$Ag_force)

#set the correct type to the different column
data$Ag_force <- as.numeric(data$Ag_force)

#The union-shop is already described as dummy, but we are going to check the class and change it to  factor
class(data$Union_shop_dummy)
#data$Union_shop_dummy <- as.factor(data$Union_shop_dummy) RETURN FALSE VALUES
```

## Exploratory Data Analysis:
Before performing any analysis on the data set, the first step is to further look at the individual variable, examine their distribution, search for any outliers that can influence and biais our model and eventually our conclusions.

```{r summary, echo = FALSE}
#Je sais pas pourquoi ca marche pas quand je knit

summary(data, digits = 2) %>%
  kable(col.names = c("State postal abbreviation", "Degree of conflict", "Degree of unionization", "Union shop dummy", "Sector", "Ag force"), caption = "Summary of the variable factors") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),font_size = 13) 
```
```{r, echo = FALSE, eval=FALSE}
#check if there are missing values:
is.null(data)
```

```{r univariate analysis, fig.pos='h', out.width= "50%", fig.show="hold", echo=FALSE}
#Let's visualize the distributions in the different columns
par(mfrow=c(1,2))

#according to the different scales it is more appropriate to separate the first numerical column from the others.
boxplot(data[,c(3,5,6)], xlab = "Variables", main = "Boxplots of numerical explanatory varibles", ylab = "Values", names=c("Degree of Unionization","Sector","Ag Force")) 

#Boxplot of the outcome variable
boxplot(data[,2], ylab = "Values", main = "Boxplot of the degree of conflict", xlab = "Degree of conflict")

```

Let's now have a look at the output variable, namely the degree of conflict.

```{r EDA ouput, echo=FALSE}
#See if it is skewed or short-/long- tails
hist(data$Degree_of_conflict, xlab = "Degree of conflict", ylab = "Density", main = "Barplot of the degree of conflict", breaks = 13, freq = FALSE)

par(new=TRUE)
plot(density(data$Degree_of_conflict), xlab = "", ylab = "", main = "", yaxt="n", xlim = c(0,4))
```
Given this figure, we can now affirm that our output follows a gaussian distribution.

Finally, concerning the dummy variable, a barplot is preferred to assess the distribution
```{r dummy, echo=FALSE}
barplot(prop.table(table(data$Union_shop_dummy)), xlab = "Union shop status", ylab = "Frequencies", main = "Distribution of the union shop status")
```
This latter figure ensures that there is fairly the same of number of observation in the two groups that we will compare in the rest of the anaysis. 


Now we will analysis the distribution of the differents variable factors according to our groups namely Right-to-work states and Union-shop States (ca on maintien que c'est trop) PAS FORCEMENT NECESSAIRE JE TROUVE QUE C?EST UN PEU OVERKILL

```{r}
#install.packages('prrr')
library('purrr')
split_sum <- data %>% split(.$Union_shop_dummy) %>% map(summary) 
knitr::kable(x = split_sum, 
      col.names = c("State postal abbreviation", "Degree of conflict", "Degree of unionization", "Union shop dummy", "Sector", "Ag force"),
      caption = "Summary of the variable factors", booktabs = TRUE)

```

Let's see how the variables are linked to each other before beginning the model fitting. It could help us make sure that the model fits properly afterwards. 

```{r, echo=FALSE}
#to assess the correlation between the variables (only the numerical one) we are going to plot scatter plots and the correlation coefficients associated

#function allowing to put the correlation cofficients on the plot
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = 1.1)
}


pairs(data[, 2:6], labels = c("Degree of conflict","Degree of unionization","Union Shop","Sector","Ag force"), main = "Correlation plots associated with the numerical variables with the Pearson correlation coefficient associated", cex.labels = 1.1, upper.panel = panel.cor)

```


## Statisitics and model fitting:

ECRIRE UN TRUC BIEN: 

The main goal first step in the analysis is a direct comparison of the observed levels of industrial conflict in the right-to-work stated and union-shop states for the period 1957-1962

- the measure of industrial conflict is the proportion of time lost due to work stoppages (man hours relative to total estimated working time)
```{r, echo = FALSE}
#Lets investiagte if there is a significant difference in the degree of conflit between unionized and non-unionized states

ttest <- t.test(data[data$Union_shop_dummy =='0',]$Degree_of_conflict,data[data$Union_shop_dummy =='1',]$Degree_of_conflict)

print(paste('The T value for the comparison between the degree of conflict between unionized and non-unionized states is',ttest$statistic,'with a p-value of',ttest$p.value)) 
```
PAS FORCEMENT NECESSAIRE: 
(RESULT PAPER: 
AVERAGE PER YEAR: 
right-to-work : 0.19% // non-right to work:0.24% --> significative at 5%

AVERAGE TOTAL MAN HOURS LOST: 
right to work: 267,151 hours per year // union-shop : 712,603 hours per year
--> significative 
--> causative factors: conceivably be favtor other than the existaence or non existence of a right to work law)

--> test the statements (better test t-test independant) : 
- Agriculture > industry in right to work whereas industry  > agriculture. 

" Where the relatively non-unionized agricultural sector is of major importance in the economy of a state, the theoretical level of industrial conflict should be less than where this is not the case".


- Try to explain the outcome variable using the variable factors (variance explained R^2 maybe adj-R^2) 
--> multiple linear regression (2 digits and ^ on estimates) : state how model fitted (ie, LS) CLEARLY describe how model selected define all terms

--> assumptions of the multiple linear regression:
1. A linear relationship between the dependent and independent variables
2. The independent variables are not highly correlated with each other (function VIF pour check si c'est bien bon)
3. The variance of the residuals is constant
4. Independence of observation

--> state the assumptions: (ON THE ERROR)
1. errors have mean 0
2. errors are homoscedastic (same variance) 
3. errors are uncorrelated
4. errors are normally distributed

```{r, echo = FALSE}
#linear Regression (WE KEEP THE INTERCEPT)

linear_mod <- lm(Degree_of_conflict ~ Degree_of_unionization + Union_shop_dummy + Sector + Ag_force  , data=data)  
summary(linear_mod)
```

Since the EDA allowed us to see that there are some outliers,  we will perform a Cook's plot to detect if there is any influential points.
```{r COOK,echo=FALSE}
#install.packages("olsrr")
ols_plot_cooksd_bar(linear_mod)
```

Given that results we sill remove those points to see if these indeed influence the results of our multiple linear regression
```{r}
data_clean <- data[-c(5,14, 17,46),]

linear_mod_clean <- lm(Degree_of_conflict ~ Degree_of_unionization + Union_shop_dummy + Sector + Ag_force  , data=data_clean)

summary(linear_mod_clean)$coef

```

#### Model Assessment:
- State how the model fitted (Ordinary Least square) degree_of_conflict ~ remaining variable factors
- check the assumptions --> qqplot of the residuals and residuals vs. fitted
                        --> function VIF pour check la mutlicollinearité (les variables indé sont elles vraiment inépendantes entre elles)


```{r residuals, echo=FALSE}

linear_mod.resid <- residuals(linear_mod)
linear_mod.fitted <- fitted(linear_mod) 

plot(linear_mod.fitted,linear_mod.resid, main="Evolution of the residuals with respect to fitted values", xlab = 'Fitted', ylab='Residuals')
lines(smooth.spline(linear_mod.fitted, linear_mod.resid),lty=2 , col="black")

legend(x=2, y=1.7, legend=c("Fitting curve", "data points" ), col=c("black", 'black'), box.lty=0, lty= c(2 , NA), pch = c(NA, 1))
```
Commenter le plot 


```{r qqplot, echo=FALSE}
#create Q-Q plot for residuals
qqnorm(linear_mod.resid)

#add a straight diagonal line to the plot
qqline(linear_mod.resid) 
```
commenter le plot


```{r diagnostic multicollinearity, echo=FALSE}
#library(caret)

car::vif(linear_mod)  %>%
  kbl(col.names = c("Variance Inflation Factor"),
#row.names = c("Degree of unionization", "Union shop dummy", "Sector", "Ag force"), Marche pas encore
      caption = "Analysis of the multicollinearity of the numerical explanatory variables") %>%
  kable_material(c("striped", "hover"))
```

## Conclusion

